---
title: "Extracting NetCDF data into Shapefiles"
author: "Julia Blanchard"
date: "27/01/2022"
output:
  html_document: default
  word_document: default
  pdf_document: default
---


```{r setup, include=FALSE}
#install.packages("ncdf4")
#install.packages("raster")
#install.packages("sf")
library(ncdf4)
library(raster)
library(sf)
library(dplyr)
library(ggplot2)
```


This is an example R script that may be useful for extracting data from FishMIP or ISIMIP NetCDF files into shapefiles, such as Large Marine Ecosystems (LMEs), Country EEZs, or regional seas.

The LME shapefiles were downloaded from:https://www.sciencebase.gov/catalog/item/55c77722e4b08400b1fd8244


```{r}
# read in shapefile
shape<-st_read("/Users/juliab6/Dropbox/FishMIP 2020:21/FishMIP Repos/LME66/LMEs66.shp")


```



The FishMIP modelled netcdf files were downloaded from ISIMIP data portal. (data.isimip.org). The filename has a lot of information in it! It is the netcdf of total catches (tc) from BOATS in the 3b simulation round of FishMIP, during the historical period (hist).


```{r}
# read in netcdf
boats<-brick("boats_ipsl-cm6a-lr_nobasd_historical_histsoc_default_tc_global_monthly_1950_2014.nc")

extent(boats) <- extent(shape)
crs(boats) <- crs(shape) 

```


A plot of the raster map of catches is here:


```{r}
boats
dim(boats) # dimensions x,y, time steps
plot(boats[[780]])
```

### Crop it and mask the catches to the lmes

Crop it and mask the catches to the lmes.

```{r}
temp<-crop(boats, extent(shape))
boats_lme<-mask(boats, shape)
plot(boats_lme[[780]])
```

# Cut to a single LME

```{r}
library(dplyr)
se_aus<-shape[46,] # south east Australia
temp<-crop(boats, extent(se_aus))
boats_seaus<-mask(temp, se_aus)
plot(boats_seaus[[780]])

```

### Extract time series for single LME

```{r}
ts_se_aus<-extract(boats_seaus, se_aus, fun=mean, na.rm=T, exact=FALSE, weights=FALSE,
   normalizeWeights=T, cellnumbers=FALSE, small=TRUE, df=T, layer=0, nl=780, 
   factors=FALSE, sp=FALSE)

plot(2:781,ts_se_aus[-1],ylab="tc",xlab="month")

```



### Get time series for all LMEs

```{r}
# Rasterize polygon shapefile - this only works for the first year
lme <-as(st_geometry(shape), 'Spatial')
lme <- rasterize(lme, boats,
  background = 0,
  filename = 'lme.grd',
  overwrite = TRUE)

crs(lme) <- crs(boats)
boats_lme <- crop(boats,lme)
boats_lme<-zonal(boats_lme,lme, "mean")


# check dimensions
dim(boats_lme)
plot(boats_lme[46,-1],ylab="tc",xlab="month")

# or use extract (verrrrry slow)

# ts_lme<-extract(boats, shape, fun=sum, na.rm=T, exact=FALSE, weights=FALSE,
#    normalizeWeights=T, cellnumbers=FALSE, small=TRUE, df=T, layer=0, nl=780, 
#    factors=FALSE, sp=FALSE)



```

### Re-arrange data and make a ggplot


```{r}
lme_number <- factor(boats_lme[, 1])
month <- 1:dim(boats_lme)[2]
boats_lme_ts <- data.frame(
  Month = rep(month, each = nrow(boats_lme)),
  LME = rep(lme_number,length(month)),
  tc = c(boats_lme))

boats_lme_ts<-boats_lme_ts[boats_lme_ts$Month>1 ,]

ggplot(boats_lme_ts,aes(x = Month, y = tc,col = LME)) + 
  geom_line()



```

### Resources

https://sesync-ci.github.io/raster-time-series-alaska-lesson/index.html

Faster version of extract() here:
https://cran.r-project.org/web/packages/exactextractr/exactextractr.pdf